---
title: "Enterochromaffin Cells"
author: "mtvector"
date: "March 21, 2018"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE,dev = c('png','pdf'))
```

A quick look at Jim's Data
# Loading Data

Following quality control, untrimmed paired-end reads were quantified using Salmon.

First I'll load the libraries and functions I'll need for the analysis (skip this part).

```{r loadLibraries,message=F,warning=F}
library(DESeq2)
library(EBSeq)
library(matrixStats)
library(biomaRt)
library(gplots)
library(clusterProfiler)
library(RColorBrewer)
library(tximport)
cols <-  colorRampPalette(rev(brewer.pal(11,"RdBu")))(50)

#see Github lengning
library(EACI)


median.normalize <- function(x){
  GetNormalizedMat(x,MedianNorm(x))
}

round.log <- function(s,base=2){
  round(log(s+1, base),digits = 1)
}

std.heatmap <- function(M,...){
  heatmap.2(M,Rowv = F,Colv = F,trace="none",col = cols,...)
}

rn.merge <- function(x,y,fill=0,simple.intersect=F){
  rn <- intersect(rownames(x),rownames(y))
  zerosx <- setdiff(rownames(x),rownames(y))
  zerosy <- setdiff(rownames(y),rownames(x))
  out <- cbind(x[rn,,drop=F],y[rn,,drop=F])
  if(simple.intersect){return(out)}
  else{
  if(length(zerosx)!=1  &length(zerosy)!=1){
    zx <- matrix(fill, nrow=length(zerosx), ncol =ncol(y), dimnames = list(zerosx,NULL))
    zy <- matrix(fill, nrow=length(zerosy), ncol =ncol(x), dimnames = list(zerosy,NULL))
    zx <- cbind(x[zerosx,],zx)
    zy <- cbind(zy,y[zerosy,])
  }else if(length(zerosx)==1){
    zx <- rep(fill, ncol(y))
    zy <- matrix(fill, nrow=length(zerosy), ncol =ncol(x), dimnames = list(zerosy,NULL))
    zx <- c(x[zerosx,],zx)
    zy <- cbind(zy,y[zerosy,])
  }else if(length(zerosy)==1){
    zx <- matrix(fill, nrow=length(zerosx), ncol =ncol(y), dimnames = list(zerosx,NULL))
    zy <- rep(fill, ncol(x))
    print(zx)
    print(zy)
    zx <- cbind(x[zerosx,],zx)
    zy <- c(zy,y[zerosy,])
  }
  out <- rbind(out,rbind(zx,zy))
  return(out)}
}

cor.compare <- function(x,y,min=0, varX=NULL ,interest.set =NULL, ...){
  d <- rn.compare(x,y)
  x <- d[[1]]
  y <- d[[2]]
  i = intersect(rownames(x), rownames(y))
  i = i[rowMaxs(x[i,],na.rm = T)>=min | rowMaxs(y[i,],na.rm = )>=min]
  if(!is.null(interest.set)){
    i = interest.set[interest.set%in%i]
  }
  if(!is.null(varX)){
    i = i[order(rowMeans(cbind(rowSds(x[i,]), rowSds(y[i,]))),decreasing = T)]
    i = i[1:ifelse(varX>length(i),length(i),varX)]
  }
  print("Num Genes:")
  print(length(i))
  return(cor(as.matrix(x[i,]),as.matrix(y[i,]), ...))
}

rn.compare <- function(x,y,fill=0){
  rn <- as.character(intersect(rownames(x),rownames(y)))
  zerosx <- setdiff(rownames(x),rownames(y))
  zerosy <- setdiff(rownames(y),rownames(x))
  zx <- matrix(fill, nrow=length(zerosx), ncol =ncol(y), dimnames = list(zerosx,NULL))
  zy <- matrix(fill, nrow=length(zerosy), ncol =ncol(x), dimnames = list(zerosy,NULL))
  nx <- rbind(x,zy)
  ny <- rbind(y,zx)
  return(list(nx[rownames(ny),,drop=F],ny[rownames(nx),,drop=F]))
}


hist.normalized <- function(x){
  nsamp <- dim(x)[2]
  h <- hist(x[,1], plot=FALSE)
  plot(h$mids, h$density, type="l", col=rainbow(nsamp)[1], main="",
       xlab="expression value", ylab="Proportion of molecules")
  for(i in 2:nsamp){
    h <- hist(x[,i], plot=FALSE)
    lines(h$mids, h$density, col=rainbow(nsamp)[i])
  }
}

```

Tried yuxi's counts tables and normalized data. All no luck. Seems like the low input has majorly affected sample identity.

```{r loadData}
# datapath <- "~/code/IngrahamLabData/JimGalaxy/"
# fileList <- dir(datapath,full.names = T)[-1]
# dsList <- lapply(fileList,read.csv2, sep="\t",header=T,row.names=1,stringsAsFactors=F)
# dsList <- list(dsList[[1]],Reduce(cbind, dsList[-1]))
# JimData <- rn.merge(dsList[[1]],dsList[[2]],simple.intersect = T)
# JimData <- apply(JimData,2,type.convert)
# colnames(JimData)[6:ncol(JimData)] <- sapply(strsplit(fileList[-1],"\\[|\\]"),"[",2)
# std.heatmap(cor(JimData,method = "spearman"))
# JimDataNorm <- median.normalize(JimData)
# barplot(JimDataNorm["Chga",],las=2,cex.names = .5)
# barplot(gutMatNorm["Chga",],las=2,cex.names = .5)
#std.heatmap(cor.compare(JimDataNorm,gutMatNorm,method = "spearman"))
# datapath <- "~/code/IngrahamLabData/JimGalaxy/counts"
# fileList <- dir(datapath,full.names = T)
# dsList <- lapply(fileList,read.csv2, sep="\t",header=T,row.names=1,stringsAsFactors=F)
# gutMat <- Reduce(rn.merge,dsList)
# colnames(gutMat) <- sapply(strsplit(fileList,"Counts_"),"[",2)
# gutMat <- gutMat[,colSums(gutMat)>500000]
# gutMatNorm <- median.normalize(gutMat)
#gutMat <- JimData
#gutMatNorm <- JimDataNorm
#sort(JimDataNorm[,1],decreasing = T)
#sort(gutMatNorm[,1],decreasing = T)


datapath <- "~/code/IngrahamLabData/JimSalmonOutputs"
fileList <- dir(datapath)
fileList <- paste0(datapath,"/",fileList,"/","quant.sf")

dsList <- lapply(fileList,read.csv2, sep="\t",header=T,row.names=1,stringsAsFactors=F)
allRownames <- Reduce(union,lapply(dsList,rownames))

#Use Biomart to get the
mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", dataset = "mmusculus_gene_ensembl", host="www.ensembl.org")
rnSymbol <- getBM(attributes = c("ensembl_transcript_id_version","mgi_symbol","description"),filters = c("ensembl_transcript_id_version"),values =allRownames ,mart = mart)
rnSymbolGenes <- rnSymbol[rnSymbol$mgi_symbol!="",]
rnSymbolGenes <- rnSymbol[rnSymbol$mgi_symbol!=""& !grepl("predicted gene", rnSymbol$description),]

txList <-  tximport(fileList,type="salmon",txOut = T)
gutMat <-  summarizeToGene(txList,rnSymbolGenes)$counts[-1,]

colnames(gutMat) <- gsub("_L003_Rout","",lapply(strsplit(fileList,split = "/"),"[",5))

std.heatmap(cor(gutMat,method="spearman"),cexCol=.5,cexRow=.5)

gutMatNorm <-  median.normalize(gutMat[rowMaxs(gutMat)>2,])
condits <- sapply(strsplit(colnames(gutMatNorm),"_"),function(x)x[1])

```

Principal component analysis for Jim's samples.

```{r JimPCA}

barplot(colSums(gutMat),main = "Number of reads",las=3,cex.names = .6)
barplot(MedianNorm(gutMat),main = "Size Factors",las=3,cex.names = .6)
hist.normalized(log(gutMat+1,2))
legend("topright",fill = rainbow(ncol(gutMat)),col=rainbow(ncol(gutMat)),legend = colnames(gutMat))
barplot(colSums(gutMatNorm),main = "PostNorm Reads",las=3,cex.names = .6)
hist.normalized(log(gutMatNorm+1,2))
legend("topright",fill = rainbow(ncol(gutMatNorm)),col=rainbow(ncol(gutMatNorm)),legend = colnames(gutMatNorm))

#gutMat <- gutMat[,c(-1,-3)]
#gutMatNorm <-  median.normalize(gutMat[rowMaxs(gutMat)>2,])
condits <- sapply(strsplit(colnames(gutMat),"_"),function(x)x[1])
hist.normalized(log(gutMatNorm+1,2))
legend("topright",fill = rainbow(ncol(gutMatNorm)),col=rainbow(ncol(gutMatNorm)),legend = colnames(gutMatNorm))

barplot(colMaxs(gutMatNorm),las=2,cex.names = .5,main="Max Expression Values per Sample")


gutCLN <- round.log(gutMatNorm+1,2)
gutCLN <- gutCLN[rowSds(gutCLN)>1,]
svgut <- svd((gutCLN-rowMeans(gutCLN))/rowSds(gutCLN))

heatmap.2(cor(gutMatNorm,method = "spearman"),col = cols,trace="none",cexCol = .5,cexRow = .5)

conditNums <- sapply(condits,function(x)which(x==sort(unique(condits))))
par(mar=c(5.1, 4.1, 4.1, 9.1), xpd=TRUE)
plot(svgut$v[,1:2],col=conditNums,xlab="PC1",ylab="PC2",pch=1:3)
legend("bottomright",inset = c(-.4,0),legend = unique(names(conditNums)),col=conditNums,pch = 1:3)
```

```{r EACI}

eaciout <- list()
l <- 1
eacivector <- svgut$u[,l]
names(eacivector) <- rownames(gutCLN)
eaciout[[l]] <- eacitest(eacivector,"org.Mm.eg","SYMBOL",sets = "GO")$setscores

l <- 2
eacivector <- svgut$u[,l]
names(eacivector) <- rownames(gutCLN)
eaciout[[l]] <- eacitest(eacivector,"org.Mm.eg","SYMBOL",sets = "GO")$setscores

```

Now to print the enrichment along each of these components

```{r PC1GO}
#PC1 Positive
a <- eaciout[[1]][eaciout[[1]]$set.mean>0,]
print(a[1:25,])
#PC1 Negative
a <- eaciout[[1]][eaciout[[1]]$set.mean<0,]
print(a[1:25,])


```

Interesting... Now PC2 

```{r PC2GO}
#PC2 Positive
a <- eaciout[[2]][eaciout[[2]]$set.mean>0,]
print(a[1:25,])
#PC2 Negative
a <- eaciout[[2]][eaciout[[2]]$set.mean<0,]
print(a[1:25,])
```

Just plot some of the markers we know and love

```{r CheckMarkers}

print(gutMat["Chga",])
barplot(gutMatNorm["Chga",],las=2,cex.names = .5,main = "Chga")
#barplot(gutMatNorm["Chga",grepl("GFP",colnames(gutMatNorm))],las=2,cex.names = .5,main = "Chga")

barplot(gutMatNorm["Lmx1a",],las=2,cex.names = .5,main = "Lmx1a")
print(gutMat["Tph1",])
barplot(gutMatNorm["Tph1",],las=2,cex.names = .5,main = "Tph1")
barplot(gutMatNorm["Scn1a",],las=2,cex.names = .5,main = "Scn1a")

heatmap.2(gutCLN[rownames(gutCLN)[grepl("Olfr",rownames(gutCLN))],],trace = "none",col = cols,main="Olfr Expression")

```

```{r xCell}

library(xCell)
bmnu <- gutMatNorm
rownames(bmnu) <- toupper(rownames(bmnu))
std.heatmap(xCell::xCellAnalysis(bmnu),cexRow=.45,cexCol=.5,main="xCell Scores")

std.heatmap(xCell::xCellSignifcanceBetaDist(xCell::xCellAnalysis(bmnu)),cexRow=.45,cexCol=.5,main="p Cell is not in mixture")
```

```{r LimitSets}



```


